---
title: "Indonesia's Deindustrialisation"
author: "Satrio Bagus Panuntun"
format: html
editor: visual
---

# The Sleeping Dragon that always sleeps: Indonesia Case

## Chapter 1: The Economic Growth Elsewhere

-   Explain briefly on the history, economic wise, leading up to the Asian Crisis
-   This includes economic profiles, trajectory and structural transition
-   Comparison to other Asian countries
-   Commentary by the IMF, ADB, World bank at that time
-   and then goes forward to today's economic condition
-   ask questions, for example: why are we still here, where is the promised land?

### TFP section and comparison to other peers

Highlight: 1. Indonesia TFP has loose its steam, economic growth driven by capital accumulation and labour. 2. Asian Financial crisis, had a significant impact on the structure of the economy. 3.

#### Indonesia TFP

```{r Indonesia TFP}
#| label: fig-growth-decomp-Indonesia
#| fig-width: 13
#| fig-height: 10
#| fig-dpi: 300
#| fig-cap: "Sources of Real GDP Growth — Indonesia"
#| fig-align: center
#| out-width: "100%"
#| warning: false
#| message: false
#| echo: false

solow_growth_decomp(data = prod_asean_5, country_pick = "Indonesia", start = "1970",end = "2023",return_fig = TRUE)
# Top accent line
grid.lines(x = c(0, 1), y = 1, gp = gpar(col = "#3E2F5B", lwd = 12.5))

# Corner accent square
grid.rect(x = 0, y = 1, width = 0.04, height = 0.03,
          just = c("left", "top"),
          gp = gpar(fill = "#3E2F5B", lwd = 0, col = NA))

grid.text(
  "Satrio Bagus Panuntun",
  x = 0.05,
  y = 0.03,
  just = c("left", "bottom"),
  gp = gpar (
    col = "grey50",
    fontsize = 9
  )
)

```

#### TFP across countries

```{r}
#| label: fig-growth-decomp
#| fig-width: 13
#| fig-height: 10
#| fig-dpi: 300
#| fig-cap: "Sources of Real GDP Growth — Indonesia vs ASEAN Peers"
#| fig-align: center
#| out-width: "100%"
#| warning: false
#| message: false
#| echo: false
p4

# Top accent line
grid.lines(x = c(0, 1), y = 1, gp = gpar(col = "#3E2F5B", lwd = 12.5))

# Corner accent square
grid.rect(x = 0, y = 1, width = 0.04, height = 0.03,
          just = c("left", "top"),
          gp = gpar(fill = "#3E2F5B", lwd = 0, col = NA))

grid.text(
  "Satrio Bagus Panuntun",
  x = 95.5,
  y = 0.03,
  just = c("right", "bottom"),
  gp = gpar (
    col = "grey50",
    fontsize = 9
  )
)
```

### Manufacturing is gone, long live commodity.

-   Manufacturing share down.
-   Rodrik (2013) stipulate that the manufacturing is still the main engine of growth for countries to transform the economy and escape middle-income trap. (find more similar paper about this topic and its relation to Indonesia).
-   Indonesia manufacturing share is down prematurely compared to other countries, this can be look at from when did the share went down and then where did the GDP per capita stand.
-   Import-substituion industrial policy in 1970-1990 (Manufacturing value added per capita graphs)
-   IS policy is basically to substitute imported goods by producing the goods within the country to fulfill domestic demand.
-   Turned to a more export oriented policy after Asian financial crisis where IMF bailout requirement was to lower import tariff.
-   Many argue that one the main reason of Indonesia not taking off is primarily due trade liberalization, commodity boom cycle after Asian financial crisis, and ineffective import substitution policy which have made firms that are shielded from the said policy to be complacent and thus unwilling to innovate and compete in the international market.
-   FDI investment has shifted from manufacturing to other commodities sector such mining and CPO.

Indonesian Industrial Policy Context

Indonesia began the experience rapid industrialisation following major political change and economic reforms in 1966-67 led by President Soeharto (Aswicahyono at al, 2011). During this period, the principal driver of the economy were agriculture and import substitution. In 1970's-1980s, however, Indonesia was benefited by the oil boom which has helped to prop up the economy out of significant hardship after a series of hyperinflation, stagnation and political turmoil from the previous decades. Revenues coming from oil exports were then invested into state-owned enterprises such as iron & steel, petroleum aluminium, and fertilizers. As the government aimed to boost the agriculture sector, these investment were part of the policy to support its growth by substituting its imported goods to be locally produced within to country, primarily on raw material production such as fertilizers, cement, chemicals, pulp and textiles. As a result, there was a large influx of workers into the manufacturing, slowly shifting away from agriculture sector.

At the end of the oil boom in 1983, oil prices dropped, so thus the oil export revenues, coupled with an economic slowdown in the 1982-1986 period. This has led the government to liberalized the economy which include variety of deregulation reforms, such as reductions in tariff and non-tariff measures, deregulation of foreign direct investment regulations, financial sector reforms and an effort to curb monopoly power of large companies through state-induced divestiture. Such reforms had transformed the industrial development direction , geared for foreign companies to invest and built factories, from inward-looking to export-oriented posture. In contrast, the domestic industry was directed to machinery equipment, medium-high technology goods and several other commodities such as aircraft, engines and shipping (Kuncoro, 2007).

## Talk about Asian financial crisis.

# Manufacturing VA % GDP Indonesia vs ASEAN & East Asian countries

```{r}



```

# Manufacturing intensity chart

# Medium-and-high technology manufactured export, Manufactured Export, and commodity export (Highly likely that some of the commodity export is being part of manufactured export. May need to redefine the definition of Manufactured export.)

# FDI movement across sectors

# Manufacture employment share.

### Excess Labour or Simply Mismanagement

### The Light at the of the tunnel, but how far are we? If so, what we can we do about it
