---
title: "Indonesia's Deindustrialisation"
author: "Satrio Bagus Panuntun"
format: html
editor: visual
---

# The Sleeping Dragon that always sleeps: Indonesia Case

## Chapter 1: The Economic Growth Elsewhere

-   Explain briefly on the history, economic wise, leading up to the Asian Crisis
-   This includes economic profiles, trajectory and structural transition
-   Comparison to other Asian countries
-   Commentary by the IMF, ADB, World bank at that time
-   and then goes forward to today's economic condition
-   ask questions, for example: why are we still here, where is the promised land?

### TFP section and comparison to other peers

### Manufacturing is gone, long live commodity

-   Manufacturing share down

### Excess Labour or Simply Mismanagement

### The Light at the of the tunnel, but how far are we?

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
