---
title: "Indonesia's Deindustrialisation"
author: "Satrio Bagus Panuntun"
format: html
editor: visual
---

# The Sleeping Dragon that always sleeps: Indonesia Case

## Chapter 1: The Economic Growth Elsewhere

-   Explain briefly on the history, economic wise, leading up to the Asian Crisis
-   This includes economic profiles, trajectory and structural transition
-   Comparison to other Asian countries
-   Commentary by the IMF, ADB, World bank at that time
-   and then goes forward to today's economic condition
-   ask questions, for example: why are we still here, where is the promised land?

### TFP section and comparison to other peers

Highlight:
1. Indonesia TFP has loose its steam, economic growth driven
by capital accumulation and labour. 
2. Asian Financial crisis, had a significant impact on the structure of the economy. 
3. 

### Manufacturing is gone, long live commodity

-   Manufacturing share down

### Excess Labour or Simply Mismanagement

### The Light at the of the tunnel, but how far are we?

#### Indonesia TFP

```{r Indonesia TFP}
#| label: fig-growth-decomp-Indonesia
#| fig-width: 13
#| fig-height: 10
#| fig-dpi: 300
#| fig-cap: "Sources of Real GDP Growth — Indonesia"
#| fig-align: center
#| out-width: "100%"
#| warning: false
#| message: false
#| echo: false

solow_growth_decomp(data = prod_asean_5, country_pick = "Indonesia", start = "1970",end = "2023",return_fig = TRUE)

```

#### TFP across countries

```{r}
#| label: fig-growth-decomp
#| fig-width: 13
#| fig-height: 10
#| fig-dpi: 300
#| fig-cap: "Sources of Real GDP Growth — Indonesia vs ASEAN Peers"
#| fig-align: center
#| out-width: "100%"
#| warning: false
#| message: false
#| echo: false
p4
```

