---
title: "Indonesia's Deindustrialisation"
author: "Satrio Bagus Panuntun"
format: html
editor: visual
---

# The Sleeping Dragon that always sleeps: Indonesia Case

## Chapter 1: The Economic Growth Elsewhere

-   Explain briefly on the history, economic wise, leading up to the Asian Crisis
-   This includes economic profiles, trajectory and structural transition
-   Comparison to other Asian countries
-   Commentary by the IMF, ADB, World bank at that time
-   and then goes forward to today's economic condition
-   ask questions, for example: why are we still here, where is the promised land?

### TFP section and comparison to other peers


#### Indonesia TFP

```{r Indonesia TFP}
#| label: fig-growth-decomp-Indonesia
#| fig-width: 13
#| fig-height: 10
#| fig-dpi: 300
#| fig-cap: "Sources of Real GDP Growth — Indonesia"
#| fig-align: center
#| out-width: "100%"
#| warning: false
#| message: false
#| echo: false

solow_growth_decomp(data = prod_asean_5, country_pick = "Indonesia", start = "1970",end = "2023",return_fig = TRUE)
# Top accent line
grid.lines(x = c(0, 1), y = 1, gp = gpar(col = "#3E2F5B", lwd = 12.5))

# Corner accent square
grid.rect(x = 0, y = 1, width = 0.04, height = 0.03,
          just = c("left", "top"),
          gp = gpar(fill = "#3E2F5B", lwd = 0, col = NA))

grid.text(
  "Satrio Bagus Panuntun",
  x = 0.95,
  y = 0.03,
  just = c("right", "bottom"),
  gp = gpar (
    col = "grey50",
    fontsize = 9
  )
)

```

#### TFP across countries
```{r}
#| label: fig-growth-decomp
#| fig-width: 13
#| fig-height: 10
#| fig-dpi: 300
#| fig-cap: "Sources of Real GDP Growth — Indonesia vs ASEAN Peers"
#| fig-align: center
#| out-width: "100%"
#| warning: false
#| message: false
#| echo: false
p4

# Top accent line
grid.lines(x = c(0, 1), y = 1, gp = gpar(col = "#3E2F5B", lwd = 12.5))

# Corner accent square
grid.rect(x = 0, y = 1, width = 0.04, height = 0.03,
          just = c("left", "top"),
          gp = gpar(fill = "#3E2F5B", lwd = 0, col = NA))

grid.text(
  "Satrio Bagus Panuntun",
  x = 95.5,
  y = 0.03,
  just = c("right", "bottom"),
  gp = gpar (
    col = "grey50",
    fontsize = 9
  )
)
```

### Manufacturing is gone, long live commodity.

-   Manufacturing share down.
-   Rodrik (2013) stipulate that the manufacturing is still the main engine of growth for countries to transform the economy and escape middle-income trap. (find more similar paper about this topic and its relation to Indonesia).
-   Indonesia manufacturing share is down prematurely compared to other countries, this can be look at from when did the share went down and then where did the GDP per capita stand.
-   Import-substituion industrial policy in 1970-1990 (Manufacturing value added per capita graphs)
-   IS policy is basically to substitute imported goods by producing the goods within the country to fulfill domestic demand.
-   Turned to a more export oriented policy after Asian financial crisis where IMF bailout requirement was to lower import tariff.
-   Many argue that one the main reason of Indonesia not taking off is primarily due trade liberalization, commodity boom cycle after Asian financial crisis, and ineffective import substitution policy which have made firms that are shielded from the said policy to be complacent and thus unwilling to innovate and compete in the international market.
-   FDI investment has shifted from manufacturing to other commodities sector such mining and CPO.

Indonesian Industrial Policy Context

Indonesia began the experience rapid industrialisation following major political change and economic reforms in 1966-67 led by President Soeharto (Aswicahyono at al, 2011). During this period, the principal driver of the economy were agriculture and import substitution. In 1970's-1980s, however, Indonesia was benefited by the oil boom which has helped to prop up the economy out of significant hardship after a series of hyperinflation, stagnation and political turmoil from the previous decades. Revenues coming from oil exports were then invested into state-owned enterprises such as iron & steel, petroleum aluminium, and fertilizers. As the government aimed to boost the agriculture sector, these investment were part of the policy to support its growth by substituting its imported goods to be locally produced within to country, primarily on raw material production such as fertilizers, cement, chemicals, pulp and textiles. As a result, there was a large influx of workers into the manufacturing, slowly shifting away from agriculture sector.

At the end of the oil boom in 1983, oil prices dropped, so thus the oil export revenues, coupled with an economic slowdown in the 1982-1986 period. This has led the government to liberalized the economy which include variety of deregulation reforms, such as reductions in tariff and non-tariff measures, deregulation of foreign direct investment regulations, financial sector reforms and an effort to curb monopoly power of large companies through state-induced divestiture. Such reforms had transformed the industrial development direction , geared for foreign companies to invest and built factories, from inward-looking to export-oriented posture. In contrast, the domestic industry was directed to machinery equipment, medium-high technology goods and several other commodities such as aircraft, engines and shipping (Kuncoro, 2007).

During the reign of the authoritarian new order regime (1967-1998), Indonesia has experienced structural changes in the economy from agriculture to manufacturing based, which can be seen from the rise of Manufcaturing share in GDP and manufacturing export.However, the asian financial crisis in 1997 has had a massive impact both economically and politically, thus changing the prior trajectory of leaping up the economic ladder. Following the 1997 crisis, Indonesia experienced significant capital flight from 1998 until 2001 (world bank 2018). Given the fact that many of these factories or firms relied on foreign debt and foreign exchange debt, the manufacturing sector was hit the hardest from the crisis. The dependency of foreign debt came from the previous policy measures to boost manufacturing sector through inviting foreign companies to off shore their factories to Indonesia, benefiting from low cost labour. On top of that, the nature of capital also changed during this period, the downfall of USSR which ended the cold war has led investment orientation to be less toward strategic/security alliance and more towards private profit, resulting capital to be mobile than ever before. Thus it becomes easier for capital to move across countries and sectors. In response, the government established new revitaliastion plan by adopting export-led recovery or expanding trade surplus to regain from the crisis (Jacob, 2004).

In the post-crisis period, from 2000s to 2010's, Indonesian economy rebounded by about 5.1% average annual growth. However, the economic structure was shifting from manufacturing to other sectors such as service and informal sectors. The shift, however, can be attributed to two key arguments: First, during this period, the rise of China complicates the revival of manufacturing in Indonesia as Chinese economic model relies heavily on Industry and exports of manufacturing goods. This resulted a loss of competitive advantage for Indonesia in Manufacturing arena and shifted to commodity based export as commodities supplying to China becomes much more lucrative than manufacturing goods within the country. Second, commodity booms have moved investment away from manufacturing to mining, oil and gas and other commodities such Palm Oil. Primary exports surge during this period by xxx %, while secondary exports gradually fell inline with the fall of manufacturing share of GDP and manufacturing export share. 


# Manufacturing VA % GDP Indonesia vs ASEAN & East Asian countries

```{r}
#| label: fig-growth-decomp
#| fig-width: 13
#| fig-height: 10
#| fig-dpi: 300
#| fig-cap: "Sources of Real GDP Growth — Indonesia vs ASEAN Peers"
#| fig-align: center
#| out-width: "100%"
#| warning: false
#| message: false
#| echo: false

p5

# Top accent line
grid.lines(x = c(0, 1), y = 1, gp = gpar(col = "#3E2F5B", lwd = 12.5))

# Corner accent square
grid.rect(x = 0, y = 1, width = 0.04, height = 0.03,
          just = c("left", "top"),
          gp = gpar(fill = "#3E2F5B", lwd = 0, col = NA))

grid.text(
  "Satrio Bagus Panuntun",
  x = 0.95,
  y = 0.03,
  just = c("right", "bottom"),
  gp = gpar (
    col = "grey50",
    fontsize = 9
  )
)


```

# Medium-and-high technology manufactured export, Manufactured Export, and commodity export (Highly likely that some of the commodity export is being part of manufactured export. May need to redefine the definition of Manufactured export.)

# Primary vs Secondary Export

# Manufacturing export intensity chart

# FDI movement across sectors

# Manufacture employment share.

### Excess Labour or Simply Mismanagement

### The Light at the of the tunnel, but how far are we? If so, what we can we do about it
